---
title: "CMTH642 Assignment 3"
author: "Garrett Mangulins - 500925990"
date: "October 30, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###
Preperation:
The dataset is used in association with the 2018 Autumn Capstone Course CKME136.

###
Importing the data set
```{r}
df <- read.csv("C:\\Users\\Garrett Mangulins\\Documents\\MCI_2014_to_2017.csv", header=TRUE, sep=",")
View(df)
```

##########
QUESTION 1
##########
Check data characteristics. Is there missing data?
```{r}
str(df)
```
The data is separated into "int", "num" and "factor" type variables

I don't need all of the columns for my Prediction Model. Here I drop irrelevent attributes:

```{r}
df.clean <- df[ , which(names(df) %in% c("premisetype", "offence", "reportedyear", "reportedmonth", "reportedday", "occurrenceyear", "occurrencemonth", "occurenceday", "MCI", "Hood_ID", "Neighbourhood"))]

View(df.clean)
```


Now to check if there is any missing data:
(I wrote this function during the course portion of the Big Data Certificate to go through a data set and check for NULL values)
```{r}
Check.Missing <- function(mydata){
  check <- sum(is.na(mydata)) #I create a sum if there is a missing value present in the data
  if(check == 0) {
    print("There are no missing values in the data set provided.")
  }
  else{
    paste0("There are: ", check, " missing values in the data set. The rows with missing values can be found here", mydata, ".Missing")
    df.Missing <- df[rowSums(is.na(mydata)) > 0,]
    View(df.Missing)
  }
} # Function I wrote in 123 to check a dataframe for NA values

Check.Missing(df.clean)
```

#######
It will be important going forward figuring out how exaclty I will deal with the data that is missing. I could either just not include the data, as it is a very small subset of the dataset, or I could find a way to impute data into it/guess the values.
My plan at the moment is to remove these rows from the data set.

*** Need to write a script to remove the rows with the same indicies as the rows with NULL values.
As of right now Iwrote in the missing rows by hand, but making this automated will result in a better project overall.

```{r}
df.cleaner <- df.clean[-c(1063,1064,1065,1108,1131,1132,1196,1197,1198,1199,1257,1258,1259,1260,1343,1344,1345,1403,1425,1742,1743,1744,1745,1950,1951,1952,1953,1954,3839,3840,11889,11890,11891,15880,15881,15882,15883,110115,110126,113749),]
```
Check to see if I removed all the rows with MULL values:
```{r}
Check.Missing(df.cleaner)
```
The message that there are no NULL values present was displayed.


**Here I am just looking into different packages R provides for predictive modelling.

I want to be able to predict the MCI (Major Crime Indicator) based on the Occurance/Reported Year/Month/Day as well as with the Neighbourhood, the Premise Type
```{r}
#install.packages("randomForest")
suppressmessage(library(randomForest))

model <- randomForest(y=as.factor(df.cleaner$MCI),x=cbind(as.factor(df.cleaner$premisetype),as.factor(df.cleaner$Neighbourhood)),ntree=10)


plot(model)
#Here I plot the confussion matrix of the regression
model$confusion
```



This is a Decision Tree Package I found online:
```{r}
#install.packages("rpart")

library(rpart)

#Create a Decision Tree from specific attributes
fit <- rpart(MCI~occurrencemonth+premisetype,df.cleaner)

#View the classification results
printcp(fit)
#View cross-validation results
plotcp(fit)

#Plot the tree
plot(fit, uniform=TRUE,
     main="Classification Tree for MCI")
text(fit, use.n=TRUE, all=TRUE, cex=.6)
```
Decision Trees do not appear to be the best choice of prediction model for this dataset. A Decision tree works best when many of the variables are binary in their choices.




K-Nearest Neighbours

*** The struggle with KNN is that the attributes have to be scalar in value. This means I have to assign a number to each of the attributes in a category.
This si not that difficult, as MCI only contians 4 different categories, months can be represented as numbers, and premisetype and offence can also be described as a number.

```{r}
#install.packages("data.table")
library(data.table)

df.cleaner$MCIfactor <- ifelse(df.cleaner$MCI %like% "Assault",1,
                                       ifelse(df.cleaner$MCI %like% "Break and Enter", 2,
                                              ifelse(df.cleaner$MCI %like% "Robbery",3,
                                                     ifelse(df.cleaner$MCI %like% "Auto Theft",4,
                                                            ifelse(df.cleaner$MCI %like% "Theft Over",5,0
                                                                   )))))

df.cleaner$premisetypefactor <- ifelse(df.cleaner$premisetype %like% "House",1,
                                       ifelse(df.cleaner$premisetype %like% "Apartment", 2,
                                              ifelse(df.cleaner$premisetype %like% "Commercial",3,
                                                     ifelse(df.cleaner$premisetype %like% "Outside",4,
                                                            ifelse(df.cleaner$premisetype %like% "Other",5,0
                                                                   )))))

df.cleaner$monthfactor <- ifelse(df.cleaner$occurrencemonth %like% "January",1,
                                       ifelse(df.cleaner$occurrencemonth %like% "February", 2,
                                        ifelse(df.cleaner$occurrencemonth %like% "March",3,
                                         ifelse(df.cleaner$occurrencemonth %like% "April",4,
                                          ifelse(df.cleaner$occurrencemonth %like% "May",5,
                                           ifelse(df.cleaner$occurrencemonth %like% "June",6,
                                            ifelse(df.cleaner$occurrencemonth %like% "July",7,                                                                ifelse(df.cleaner$occurrencemonth %like% "August",8,
                                            ifelse(df.cleaner$occurrencemonth %like% "September",9,             
                                              ifelse(df.cleaner$occurrencemonth %like% "October",10,
                                              ifelse(df.cleaner$occurrencemonth %like% "November",11,
                                              ifelse(df.cleaner$occurrencemonth %like% "December",12,0))))))))))))

View(df.cleaner)
```

*** Time permitting, look into cleaner way to make factors for the categorical values.
 

Now that I have my categorical data as Factors
```{r}

#First I want to take only the factor columns

df.factors <- df.cleaner[,c(8,11,12,13)]

# Here I create a random sample from the normalized dataset. I am making a 70/30 split for the training and testing sets
df.split <- sample(nrow(df.factors), floor(nrow(df.factors)*0.7))

## Need to make the attributes here numeric.
## Use as.Numeric to do so

df.train <- df.factors[df.split,]
df.test <- df.factors[-df.split,]

View(df.train)

df.train.labels <- df.train[,1]
df.test.labels <- df.test[,1]

Check.Missing(df.split)
```

```{r}
#install.packages("class")
#install.packages("gmodels")
library(class)
library(gmodels)
# First create a prediction model using the knn function. As of right now it does not work.
prediction.test <- knn(train=df.train, test=df.test, cl=df.train.labels, k=10)

# Next we produce a table which shows how the 
CrossTable(x=df.test.labels, y=prediction.test, prop.chisq=FALSE)
```

